Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2024-12-09 transient.tex
Draft for committee before defense

Section \ref{sec:rod-drop} covers the model setup, numerical results, and discussion of the rod
drop experiment. Section \ref{sec:reactivity-insertion} covers the model setup, numerical results,
and discussion of the reactivity insertion experiment. Lastly, Section \ref{sec:transient-summary}
summarizes the goals and outcomes in this chapter.

@@

This work aims to replicate the integral count curve following a rod drop of Rod 1 with the
$^{235}$U loading corresponding to ``30 capsule additions'' of highly-enriched $^{235}$U capsules
into the ``pump bowl''. Each capsule contains 85 g of $^{235}$U, of which 94.94 \% enters the core
due to residual fuel salt remaining in the drain tanks after mixing \cite{fratoni_molten_2020}.

@@

\begin{figure}[htb!]
  \centering
%  \includegraphics[width=0.8\columnwidth]{rod-height}
  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion, and \gls{DNP}
  solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
  \label{fig:rod-drop-coupling}
\end{figure}

@@

Figure \ref{fig:integral-count} compares the integral neutron count during the rod drop experiment
from \gls{MSRE} experimental data and the hybrid method numerical results. The hybrid method
reproduces the expected shape of the integral
neutron count characterized by the drop in neutron count rate leading to a gradual flattening of
the integral count curve. The discrepancy in integral neutron count at around $t=16$ s is x \%.
Experimental uncertainties surrounding the rod drop experiment include the actual $^{235}$U
loading, initial rod height, rod acceleration, and the initial neutron count rate. These uncertainties
may have affected the reactivity withdrawal estimates in Section \ref{sec:rod-drop-setup}.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{integral-count}
  \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
  and hybrid method numerical results. The hybrid method agrees closely with the \gls{MSRE} data
  with a discrepancy of x \% at around $t=16$ s.}
  \label{fig:integral-count}
\end{figure}

Figures * show snapshots of the neutron group flux distributions at various times during the rod
drop simulation. The neutron flux magnitudes fall as the reactor enters a sub-critical state. The
Rod 1 induces a significant drop in neutron fluxes near the center of the core.

\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}

After a long operational run on $^{235}$U fuel, \gls{ORNL} extended the \gls{MSRE} program to
demonstrate reactor operation with $^{233}$U fuel. The \gls{MSRE} then became the first nuclear
reactor to run on $^{233}$U bred from $^{232}$Th. During a series of tests for experimental
dynamic analyses, \gls{MSRE} researchers conducted reactivity insertion tests with the reactor
operating at 1, 5, and 8 MW \cite{steffy_experimental_1970}.
In contrast to the rod drop experiments, these reactivity insertion
experiments were conducted with the \gls{MSRE} in power operation.

This work aims to replicate the power response following a reactivity insertion of 24.8 pcm of the
\gls{MSRE} initially at 1 MW. Figure \ref{fig:msre-reactivity-insertion} shows the theoretical and
experimental power response following the reactivity insertion. Three separate publications
\cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022} noted that the plot is
mislabeled with the wrong reactivity insertion value of 0.0139 \% $\delta k/k$, which had been the
reactivity insertion value for the 8 MW case. The correct reactivity insertion value is 24.8 pcm as
mentioned prior.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{msre-reactivity-insertion}
  \caption{Power response following a step reactivity insertion of 24.8 pcm with the reactor
    initially at 1 MW \cite{steffy_experimental_1970}. Note that the authors of the \gls{MSRE}
    report mislabeled the reactivity insertion value on the plot
  \cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022}.}
  \label{fig:msre-reactivity-insertion}
\end{figure}

\subsection{Model Setup}

For the reactivity insertion simulation, we must first identify the $^{233}$U salt composition used
in the experiment. Power operations with the $^{233}$U-fueled \gls{MSRE} core
occurred during Runs 16 to 18 according
to \gls{MSRE} reports \cite{thoma_chemical_1971, burke_analysis_2019}. Table \ref{table:u233-salt}
lists the proportion of elements in the \gls{MSRE} fuel salt during Runs 16 to 18. Note that the
components do not sum up to 100 \% due to experimental uncertainty in chemical analysis and the
presence of other elements that were not measured. Taking the nominal uranium value to be true, I scaled
up the compositions of non-uranium elements to make up the remaining $(100\%-0.8155\%)$ of the salt. I
kept the nominal weight percentage of uranium constant because it is the most accurate value in the table
and has the biggest impact on the reactivity.
%
\begin{table}[t]
  \small
  \centering
  \setlength\tabcolsep{4pt}
  \caption{\gls{MSRE} fuel salt mean composition during the power operations with $^{233}$U \cite{thoma_chemical_1971}.
  All values were measured via chemical analysis except the nominal U weight percentage, which was
  calculated from inventory tracking of cumulative U additions to the salt.}
  \begin{tabular}{l S[table-format=2.2(2)] S[table-format=1.2(2)] S[table-format=2.2(2)] S[table-format=2.2(3)] S[table-format=1.3(2)] S}
    \toprule
    Element & {Li} & {Be} & {Zr} & {F} & {U} & {U (Nominal)} \\
    \midrule
    Weight percentage [\%]  & 11.33(39) & 6.71(30) & 10.99(29) & 68.71(126) & 0.809(11) & 0.8155 \\
    \bottomrule
  \end{tabular}
  \begin{tabular}{l S[table-format=2(1)] S[table-format=3(2)] S[table-format=2(2)] S}
    \toprule
    Element & {Cr} & {Fe} & {Ni} & {O} \\
    \midrule
    Parts per million [ppm] & 72(9) & 135(24) & 46(12) & 61 \\
    \bottomrule
  \end{tabular}
  \label{table:u233-salt}
\end{table}

For the isotopic content of uranium, I used a composition restruction strategy developed by Burke
\cite{burke_analysis_2019}. Their analysis determined that the $^{233}$U salt consists of residual
salt from $^{235}$U operations, a depleted uranium feed, and a stock $^{233}$U feed. The residue
and depleted uranium feed masses, taken from Burke's work, are 1.935 kg and 0.890 kg, respectively.
The stock feed uranium
mass is estimated to be 35.234 kg reported on January 12 1969 \cite{thoma_chemical_1971}. The
experiment was performed on January 16 1969 \cite{steffy_experimental_1970}. Table * shows
the breakdown and total isotopic content of uranium in the salt. The density of $^{233}$U salt is
2.264 g cm$^{-3}$ at 911 K \cite{thoma_chemical_1971}.
%
\begin{table}[t]
  \small
  \centering
  \setlength\tabcolsep{4pt}
  \caption{Uranium isotopic composition reconstruction table \cite{burke_analysis_2019}.}
  \begin{tabular}{l S S S S S S S}
    \toprule
    \multirow{2}{*}{Uranium Source} & {\multirow{2}{*}{Mass [kg]}} & \multicolumn{6}{c}{Weight percentage [\%]} \\
                                    & & {$^{232}$U} & {$^{233}$U} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} & {$^{238}$U} \\
                                    \cmidrule(r){1-1} \cmidrule(rl){2-2} \cmidrule(l){3-8}
    Residue & 1.935 & & & 0.343 & 32.705 & 0.483 & 66.469 \\
    Depleted & 0.890 & & & & 0.3 & & 99.7 \\
    Stock & 35.234 & 0.02 & 91.49 & 7.6 & 0.7 & 0.05 & 0.14 \\
    \bottomrule
  \end{tabular}
  \label{table:u233-composition}
\end{table}

Due to constraints on computational resources, the reactivity insertion simulation ran a 3-D
quarter-core \gls{MSRE} model. Refer to Figure \ref{fig:1/4-geom} for the horizontal cross section
view of the 3-D quarter-core model. Figure * shows the side view from a vertical side of the
quarter-core and the direction of salt flow in each major region. The regions are the salt-graphite
lattice, the upper plenum, the out-of-core components, the downcomer, and the lower plenum. The
model assumes a flat salt velocity profile in each region. The actual total volumetric flow rate of
the \gls{MSRE} is 1200 gpm (gallons per minute) \cite{robertson_msre_1965}. Dividing this value by
four for the quarter-core model and converting to SI units gives a volumetric flow rate of
0.01893 m$^3$ s$^{-1}$.

Table \ref{table:msre-flow} lists the dimensions, salt flow velocities, and salt residence
times in each region to maintain the calculated volumetric flow rate. The salt residence times are
consistent with \gls{MSRE} specifications \cite{robertson_msre_1965}. The out-of-core region is
modeled as a 1-D pipe due to its lower neutronic importance other than \gls{DNP} decay and heat
removal through the heat exchanger. The out-of-core region salt velocity is set to the same magnitude
as the downcomer for ease of implementation with the height adjusted accordingly to maintain the
out-of-core salt residence time. Note that the velocity in the downcomer is negative since salt
flows downwards unlike in the other regions.
%
\begin{table}[t]
  \small
  \centering
  \caption{Dimensions, salt flow velocities, and salt residence times in each region of the \gls{MSRE}.}
  \begin{tabular}{l S S S S}
    \toprule
    Component & {Salt cross-sectional area [m$^2$]} & {Height [m]} & {Velocity [m s$^{-1}$]} & {Time [s]} \\
    \cmidrule(r){1-1} \cmidrule(l){2-5}
    Upper plenum & 0.42411 & 0.2539 & 0.04463 & 5.69 \\
    Lattice      & 0.08646 & 1.6637 & 0.21891 & 7.60 \\
    Lower plenum & 0.42614 & 0.1875 & 0.04935 & 3.80 \\
    Out-of-core  & {-}     & 5.3081 & 0.65532 & 8.1 \\
    Downcomer    & 0.02888 & 1.6637 & -0.65532 & 2.539 \\
    \bottomrule
  \end{tabular}
  \label{table:msre-flow}
\end{table}

% Unlike with $^{235}$U salt, there is no widely recognized benchmark value
% for the density of $^{233}$U salt. During \gls{MSRE} operation, researchers were still developing
% reliable techniques for measuring the density of molten salt mixtures \cite{thoma_chemical_1971}.
% \gls{MSRE} $^{233}$U fuel salt with has a lower density than $^{235}$U salt largely due to its
% lower $^{238}$U content.

The MultiApp nested iterative coupling structure for this reactivity insertion simulation is similar
to the implementation for the rod drop simulation. Figure \ref{fig:insertion-coupling}
provides an updated illustration of the nexted iteration coupling with temperature. The \gls{DNP}
and temperature distributions are solved together with additive Schwarz method preconditioning in
Moltres since they are both advective problems. The 1-D out-of-core model accounts for \gls{DNP}
loss and heat removal outside the core. Delayed neutrons produced in this region are considered
lost and do not contribute to the neutron flux in the core.

\begin{figure}[htb!]
  \centering
%  \includegraphics[width=0.8\columnwidth]{rod-height}
  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
  \gls{DNP}-temperature, and out-of-core
  solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
  \label{fig:insertion-coupling}
\end{figure}

The reactivity insertion simulation requires a prior ``steady-state'' simulation to obtain the group
flux, \gls{DNP}, and temperature distributions under 1 MW power operation and steady salt flow. The
``steady-state'' simulation is a time-dependent simulation with timestep sizes of 1 s to advance
the model towards 1 MW steady state operation. Following this, the reactivity insertion simulation
uses the \gls{MOOSE} Restart capability to load in variable distributions for the initial conditions.

\subsection{Preliminary Rod Worth Calculations}

The control rod worth of the \gls{MSRE} quarter-core model with $^{233}$U fuel must be determined
to estimate the initial rod height and rod travel necessary for the reactivity insertion
simulation. We start this process with comparisons of Rod 1 worths between a full-core OpenMC model
and experimental data. Table \ref{table:u233-rod-worth} compares the total rod worth of Rod 1 at
922 K between the OpenMC model from this work
and \gls{MSRE} data. The rod worth estimate from OpenMC agrees closely with the predicted (i.e.,
calculated) \gls{MSRE} value. The measured \gls{MSRE} rod worth value is about 7 \% smaller than
the predicted value. \gls{MSRE} researchers noted poor precision in their rod worth measurements
due to ``wear and aging'' in the control rod drive assembly \cite{engel_zero-power_1972}. This may
be responsible for some uncertainty. However, the actual rod worth is likely closer to the measured
value than the predicted value, and the \gls{MSRE} model of this work is likely to contain as yet
unknown biases. Nevertheless, the rod worth estimate is close enough that it should not affect the
\gls{MSRE} power response following a fixed reactivity insertion.
%
\begin{table}[t]
  \centering
  \caption{Total rod worth of Rod 1 at 922 K from OpenMC full-core \gls{MSRE} model and \gls{MSRE}
  experimental data.}
  \begin{tabular}{l S[table-format=4(3)] S}
    \toprule
    & {Rod worth [pcm]} & {Percentage discrepancy [\%]} \\
    \cmidrule(r){1-1} \cmidrule(l){2-3}
    MSRE (measured) & 2580(129) & {-} \\
    MSRE (predicted) & 2750 & 6.6 \\
    OpenMC (this work) & 2771 & 7.4\\
    \bottomrule
  \end{tabular}
  \label{table:u233-rod-worth}
\end{table}

The 3-D quarter-core \gls{MSRE} model for the reactivity insertion simulation contains only one
control rod, but it is equivalent to four rods due to rotational symmetry of the quarter-core model.
The total rod worth of the 3-D quarter-core \gls{MSRE} model using the hybrid method on Moltres is
9702 pcm, just under four times the worth of Rod 1 from Table \ref{table:u233-rod-worth} due to
rod shadowing effects that reduce the worth of adjacent control rods. Figure \ref{fig:u233-rod-worth-curve} shows the rod
worth values from Moltres for the 3-D quarter-core model and the monotone cubic spline interpolation
curve. Despite having fewer data points than the $^{235}$U rod worth analysis (Section
\ref{sec:3d-nts}), this interpolation curve is valid for the reactivity insertion experiment since
the rod heights will strictly remain within the middle
range where the change in rod worth is approximately linear to changes in rod height.

The required
rod travel in the quarter-core model for a 24.8 pcm reactivity insertion in this linear region
is 0.105 inches (0.267 cm). The estimated rod travel required to induce the same reactivity
insertion in the actual \gls{MSRE} reactor is 0.396 inches (1.00 cm) obtained by scaling with the
total measured \gls{MSRE} rod worth in Table \ref{table:u233-rod-worth}. Using
an estimated rod speed of 1.27 cm s$^{-1}$ \cite{robertson_msre_1965}, the reactivity insertion
occurs over a 0.791-s ramp (reactivity insertions at this timescale are considered step reactivity
insertions). The rod is assumed to accelerate and decelerate instantaneously given
the low rod speed.
The reactivity insertion simulation ran with timestep sizes of 0.198 s during rod movement and
0.25 s after the rod reaches its final position. These timestep sizes were chosen based on
considerations balancing accuracy and available computational resources. The main purpose of this
work is the demonstration of the hybrid method for time-dependent control rod modeling; a detailed
time-stepping truncation error analysis is out of the scope of this work.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
  \caption{Rod worth data points and the monotone cubic spline interpolation curve from $k$-eigenvalue
  hybrid method simulations on the 3-D quarter-core \gls{MSRE} model.}
  \label{fig:u233-rod-worth-curve}
\end{figure}

\subsection{Reactivity Insertion Numerical Results \& Discussion}

Figure \ref{fig:insertion-power} shows the \gls{MSRE} data and simulation results for change in
power output following the 24.8 pcm reactivity
insertion. From the experimental data, the power output rises rapidly at the start of the
reactivity insertion before oscillating back down below the initial power output. Power output
peaks again later at around $t=360$ s. Moltres reproduces the same trend in power output, but 
over/under-predicts the first power peak by y \%. It also over/under-predicts the subsequent power
oscillations.
Significant contributors to the discrepancies are the uncertainties in the physical properties of
the fuel salt, temperature reactivity coefficient, and the dynamic response of the heat exchanger.

Figure \ref{fig:insertion-temperature} shows the maximum fuel salt and graphite temperatures, and
the mean salt inlet and outlet
temperatures following the reactivity insertion. The increase in maximum temperatures is more
gradual than the increase in power output as the salt and graphite take time to heat up. The salt
outlet temperature follows closely behind as the heated salt flows out of the core. The salt inlet
temperature lags behind the other temperature measurements due to the time taken for the heated
salt to travel through the out-of-core region. The temperature values appear to settle at higher
temperatures to compensate for the inserted reactivity.

\begin{figure}[t]
  \centering
%  \includegraphics[width=0.8\columnwidth]{u233-rod-worth-curve}
  \caption{Change in power output following a step reactivity insertion of 24.8 pcm with the
  \gls{MSRE} initially at 1 MW.}
  \label{fig:insertion-power}
\end{figure}

>>>>

Section \ref{sec:rod-drop} covers the model setup, numerical results, and discussion of the rod
drop experiment.
%Section \ref{sec:reactivity-insertion} covers the model setup, numerical results,
%and discussion of the reactivity insertion experiment.
Section \ref{sec:transient-summary}
summarizes the findings from the rod drop experiment study.

@@

This work aims to replicate the integral count curve following a rod drop of Rod 1 with the
$^{235}$U loading corresponding to ``30 capsule additions'' of highly-enriched $^{235}$U capsules
into the \gls{MSRE} through the ``pump bowl''. Each capsule contains 85 g of $^{235}$U, of which
94.94 \% enters the core
due to residual fuel salt remaining in the drain tanks after mixing \cite{fratoni_molten_2020}.

@@

\begin{figure}[t]
  \tikzstyle{every node}=[font=\small]
  \centering
  \begin{tikzpicture}
    \node (1) [solver] {\textbf{$\bm{S_N}$ transport subsolver}};
    \node (2) [solver, right of=1, xshift=4cm] {\textbf{Neutron\\ diffusion\\ subsolver}};
    \node (3) [solver, right of=2, xshift=4cm] {\textbf{DNP subsolver}};
    \draw [arrow] (1.10) -- node[anchor=south, text width=4cm, align=center] {$\vec{D}_g, \gamma_g$} (2.170);
    \draw [arrow] (2.190) -- node[anchor=north, text width=4cm, align=center] {$\phi_g, C_i$} (1.350);
    \draw [arrow] (2.10) -- node[anchor=south, text width=4cm, align=center] {$Q_f, Q$} (3.170);
    \draw [arrow] (3.190) -- node[anchor=north, text width=4cm, align=center] {$C_i$} (2.350);
    \node (5) [bound, fit=(2) (3), inner sep=8pt, label={[align=left]above:\textbf{Inner iteration}}] {};
    \node (6) [bound, fit=(1) (5), inner sep=15pt, label={[align=left]above:\textbf{Outer iteration}}] {};
  \end{tikzpicture}
  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion, and \gls{DNP}
  solvers for the rod drop simulation using the hybrid $S_N$-diffusion method.}
  \label{fig:rod-drop-coupling}
\end{figure}

@@

Figure \ref{fig:count-rate} shows the neutron count rate following the rod drop from Moltres. It
shows the steep initial decline in neutron count rate as the rod drops further into the core. The
decline slows at around $t=0.5$ s due to the delayed response of \glspl{DNP}. During this
time-dependent simulation, convergence issues prevented it from converging beyond $t=0.825$ s. The
simulation managed to continue after raising the convergence tolerance value of the neutron
diffusion subsolver and the fixed point iterations from $\epsilon_\text{tol}=10^{-8}$ to $10^{-7}$. However,
this action led to under-converged scalar flux and neutron count results between $t=1.325$ s and
$3.825$ s.

Figure \ref{fig:integral-count} compares the integral neutron count during the rod drop experiment
from \gls{MSRE} experimental data and the hybrid method numerical results. The hybrid method
reproduces the expected shape of the integral
neutron count characterized by the drop in neutron count rate leading to a gradual flattening of
the integral count curve. Moltres underpredicts the integral neutron count throughout the
simulation compared to the \gls{MSRE} experimental data.
Experimental uncertainties surrounding the rod drop experiment include the actual $^{235}$U
concentration, initial \& final rod heights, rod acceleration, and the initial neutron count rate.
These uncertainties may have affected the reactivity withdrawal estimates in Section
\ref{sec:rod-drop-setup}.

Considering that the \gls{DNP} subsolver convergence remained unaffected throughout the simulation,
it would have produced \gls{DNP} distributions that are fairly accurate. Only minor discrepancies
may be present in the \gls{DNP} distributions through \gls{DNP} production from under-converged
scalar fluxes. Figure \ref{fig:count-rate-interp} shows the neutron count rate if we replaced the
under-converged data points between $t=1.325$ s and $3.825$ s with linearly interpolated count
rates from $t=0.825$ s to $4.325$ s. Figure \ref{fig:integral-count-interp} shows the corresponding
integral neutron count curve generated using the interpolated data. In this case, the integral
count curve shape is more consistent with the experimental data, albeit seemingly in response to a
larger reactivity withdrawal.

\begin{figure}[p]
  \centering
  \includegraphics[width=0.8\columnwidth]{count-rate}
  \caption{Neutron count rate during the rod drop experiment from Moltres. The convergence tolerance
  of the neutron diffusion subsolver and the fixed point iterations was raised from $\epsilon=10^{-8}$
  to $10^{-7}$ from $t=1.32$ s onwards.}
  \label{fig:count-rate}
  \includegraphics[width=0.8\columnwidth]{integral-count}
  \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
  and Moltres.}
  \label{fig:integral-count}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.8\columnwidth]{count-rate-interp}
  \caption{Neutron count rate during the rod drop experiment from Moltres with linearly interpolated
    data between $t=1.325$ s and $3.825$ s.}
  \label{fig:count-rate-interp}
  \includegraphics[width=0.8\columnwidth]{integral-count-interp}
  \caption{Integral neutron count during the rod drop experiment from \gls{MSRE} experimental data
  and Moltres with linearly interpolated count rate data between $t=1.325$ s and $3.825$ s.}
  \label{fig:integral-count-interp}
\end{figure}

Lagging solution states during the first fixed point iteration in each time-step also slowed
solution convergence, particularly when the control rod was in motion. This limited the simulation
from taking larger time-steps. Further work could include more sophiscated coupling schemes to
help the solver converge faster.

%Figures * show snapshots of the neutron group flux distributions at various times during the rod
%drop simulation. The neutron flux magnitudes fall as the reactor enters a sub-critical state. The
%Rod 1 induces a significant drop in neutron fluxes near the center of the core.

%\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}
%
%After a long operational run on $^{235}$U fuel, \gls{ORNL} extended the \gls{MSRE} program to
%demonstrate reactor operation with $^{233}$U fuel. The \gls{MSRE} then became the first nuclear
%reactor to run on $^{233}$U bred from $^{232}$Th. During a series of tests for experimental
%dynamic analyses, \gls{MSRE} researchers conducted reactivity insertion tests with the reactor
%operating at 1, 5, and 8 MW \cite{steffy_experimental_1970}.
%In contrast to the rod drop experiments, these reactivity insertion
%experiments were conducted with the \gls{MSRE} in power operation.
%
%This work aims to replicate the power response following a reactivity insertion of 24.8 pcm of the
%\gls{MSRE} initially at 1 MW. Figure \ref{fig:msre-reactivity-insertion} shows the theoretical and
%experimental power response following the reactivity insertion. Three separate publications
%\cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022} noted that the plot is
%mislabeled with the wrong reactivity insertion value of 0.0139 \% $\delta k/k$, which had been the
%reactivity insertion value for the 8 MW case. The correct reactivity insertion value is 24.8 pcm as
%mentioned prior.
%
%\begin{figure}[t]
%  \centering
%  \includegraphics[width=0.8\columnwidth]{msre-reactivity-insertion}
%  \caption{Power response following a step reactivity insertion of 24.8 pcm with the reactor
%    initially at 1 MW \cite{steffy_experimental_1970}. Note that the authors of the \gls{MSRE}
%    report mislabeled the reactivity insertion value on the plot
%  \cite{zanetti_innovative_2015, singh_study_2019, mochizuki_validation_2022}.}
%  \label{fig:msre-reactivity-insertion}
%\end{figure}
%
%\subsection{Model Setup}
%
%For the reactivity insertion simulation, we must first identify the $^{233}$U salt composition used
%in the experiment. Power operations with the $^{233}$U-fueled \gls{MSRE} core
%occurred during Runs 16 to 18 according
%to \gls{MSRE} reports \cite{thoma_chemical_1971, burke_analysis_2019}. Table \ref{table:u233-salt}
%lists the proportion of elements in the \gls{MSRE} fuel salt during Runs 16 to 18. Note that the
%components do not sum up to 100 \% due to experimental uncertainty in chemical analysis and the
%presence of other elements that were not measured. Taking the nominal uranium value to be true, I scaled
%up the compositions of non-uranium elements to make up the remaining $(100\%-0.8155\%)$ of the salt. I
%kept the nominal weight percentage of uranium constant because it is the most accurate value in the table
%and has the biggest impact on the reactivity.
%%
%\begin{table}[t]
%  \small
%  \centering
%  \setlength\tabcolsep{4pt}
%  \caption{\gls{MSRE} fuel salt mean composition during the power operations with $^{233}$U \cite{thoma_chemical_1971}.
%  All values were measured via chemical analysis except the nominal U weight percentage, which was
%  calculated from inventory tracking of cumulative U additions to the salt.}
%  \begin{tabular}{l S[table-format=2.2(2)] S[table-format=1.2(2)] S[table-format=2.2(2)] S[table-format=2.2(3)] S[table-format=1.3(2)] S}
%    \toprule
%    Element & {Li} & {Be} & {Zr} & {F} & {U} & {U (Nominal)} \\
%    \midrule
%    Weight percentage [\%]  & 11.33(39) & 6.71(30) & 10.99(29) & 68.71(126) & 0.809(11) & 0.8155 \\
%    \bottomrule
%  \end{tabular}
%  \begin{tabular}{l S[table-format=2(1)] S[table-format=3(2)] S[table-format=2(2)] S}
%    \toprule
%    Element & {Cr} & {Fe} & {Ni} & {O} \\
%    \midrule
%    Parts per million [ppm] & 72(9) & 135(24) & 46(12) & 61 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:u233-salt}
%\end{table}
%
%For the isotopic content of uranium, I used a composition restruction strategy developed by Burke
%\cite{burke_analysis_2019}. Their analysis determined that the $^{233}$U salt consists of residual
%salt from $^{235}$U operations, a depleted uranium feed, and a stock $^{233}$U feed. The residue
%and depleted uranium feed masses, taken from Burke's work, are 1.935 kg and 0.890 kg, respectively.
%The stock feed uranium
%mass is estimated to be 35.234 kg reported on January 12 1969 \cite{thoma_chemical_1971}. The
%experiment was performed on January 16 1969 \cite{steffy_experimental_1970}. Table 
%\ref{table:u233-composition} shows
%the breakdown and total isotopic content of uranium in the salt. The density of $^{233}$U salt is
%2.264 g cm$^{-3}$ at 911 K \cite{thoma_chemical_1971}.
%%
%\begin{table}[t]
%  \small
%  \centering
%  \setlength\tabcolsep{4pt}
%  \caption{Uranium isotopic composition reconstruction table \cite{burke_analysis_2019}.}
%  \begin{tabular}{l S S S S S S S}
%    \toprule
%    \multirow{2}{*}{Uranium Source} & {\multirow{2}{*}{Mass [kg]}} & \multicolumn{6}{c}{Weight percentage [\%]} \\
%                                    & & {$^{232}$U} & {$^{233}$U} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} & {$^{238}$U} \\
%                                    \cmidrule(r){1-1} \cmidrule(rl){2-2} \cmidrule(l){3-8}
%    Residue & 1.935 & & & 0.343 & 32.705 & 0.483 & 66.469 \\
%    Depleted & 0.890 & & & & 0.3 & & 99.7 \\
%    Stock & 35.234 & 0.02 & 91.49 & 7.6 & 0.7 & 0.05 & 0.14 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:u233-composition}
%\end{table}
%
%Due to constraints on computational resources, the reactivity insertion simulation ran a 3-D
%quarter-core \gls{MSRE} model. Refer to Figure \ref{fig:1/4-geom} for the horizontal cross section
%view of the 3-D quarter-core model. Figure * shows the side view from a vertical side of the
%quarter-core and the direction of salt flow in each major region. The regions are the salt-graphite
%lattice, the upper plenum, the out-of-core components, the downcomer, and the lower plenum. The
%model assumes a flat salt velocity profile in each region. The actual total volumetric flow rate of
%the \gls{MSRE} is 1200 gpm (gallons per minute) \cite{robertson_msre_1965}. Dividing this value by
%four for the quarter-core model and converting to SI units gives a volumetric flow rate of
%0.01893 m$^3$ s$^{-1}$.
%
%Table \ref{table:msre-flow} lists the dimensions, salt flow velocities, and salt residence
%times in each region to maintain the calculated volumetric flow rate. The salt residence times are
%consistent with \gls{MSRE} specifications \cite{robertson_msre_1965}. The out-of-core region is
%modeled as a 1-D pipe due to its lower neutronic importance other than \gls{DNP} decay and heat
%removal through the heat exchanger. The out-of-core region salt velocity is set to the same magnitude
%as the downcomer for ease of implementation with the height adjusted accordingly to maintain the
%out-of-core salt residence time. Note that the velocity in the downcomer is negative since salt
%flows downwards unlike in the other regions.
%%
%\begin{table}[t]
%  \small
%  \centering
%  \caption{Dimensions, salt flow velocities, and salt residence times in each region of the \gls{MSRE}.}
%  \begin{tabular}{l S S S S}
%    \toprule
%    Component & {Salt cross-sectional area [m$^2$]} & {Height [m]} & {Velocity [m s$^{-1}$]} & {Time [s]} \\
%    \cmidrule(r){1-1} \cmidrule(l){2-5}
%    Upper plenum & 0.42411 & 0.2539 & 0.04463 & 5.69 \\
%    Lattice      & 0.08646 & 1.6637 & 0.21891 & 7.60 \\
%    Lower plenum & 0.42614 & 0.1875 & 0.04935 & 3.80 \\
%    Out-of-core  & {-}     & 5.3081 & 0.65532 & 8.1 \\
%    Downcomer    & 0.02888 & 1.6637 & -0.65532 & 2.539 \\
%    \bottomrule
%  \end{tabular}
%  \label{table:msre-flow}
%\end{table}
%
%% Unlike with $^{235}$U salt, there is no widely recognized benchmark value
%% for the density of $^{233}$U salt. During \gls{MSRE} operation, researchers were still developing
%% reliable techniques for measuring the density of molten salt mixtures \cite{thoma_chemical_1971}.
%% \gls{MSRE} $^{233}$U fuel salt with has a lower density than $^{235}$U salt largely due to its
%% lower $^{238}$U content. 
%
%The MultiApp nested iterative coupling structure for this reactivity insertion simulation is similar
%to the implementation for the rod drop simulation. Figure \ref{fig:insertion-coupling}
%provides an updated illustration of the nexted iteration coupling with temperature. The \gls{DNP}
%and temperature distributions are solved together with additive Schwarz method preconditioning in
%Moltres since they are both advective problems. The 1-D out-of-core model accounts for \gls{DNP}
%loss and heat removal outside the core. Delayed neutrons produced in this region are considered
%lost and do not contribute to the neutron flux in the core.
%
%\begin{figure}[t]
%  \tikzstyle{every node}=[font=\small]
%  \centering
%  \begin{tikzpicture}
%    \node (1) [solver] {\textbf{$\bm{S_N}$ transport subsolver}};
%    \node (2) [solver, right of=1, xshift=4cm] {\textbf{Neutron\\ diffusion\\ subsolver}};
%    \node (3) [solver, right of=2, xshift=4cm] {\textbf{In-core DNP \& temperature subsolver}};
%    \node (4) [solver, below of=3, yshift=-3cm] {\textbf{Ex-core DNP \& temperature subsolver}};
%    \draw [arrow] (1.10) -- node[anchor=south, text width=4cm, align=center] {$\vec{D}_g, \gamma_g$} (2.170);
%    \draw [arrow] (2.190) -- node[anchor=north, text width=4cm, align=center] {$\phi_g, C_i, T$} (1.350);
%    \draw [arrow] (2.10) -- node[anchor=south, text width=4cm, align=center] {$Q_f, Q$} (3.170);
%    \draw [arrow] (3.190) -- node[anchor=north, text width=4cm, align=center] {$C_i, T$} (2.350);
%    \draw [arrow] (3.280) -- node[anchor=west, text width=1cm, align=center] {$C^\text{out}_i$\\ $T^\text{out}$} (4.80);
%    \draw [arrow] (4.100) -- node[anchor=east, text width=1cm, align=center] {$C^\text{in}_i$\\ $T^\text{in}$} (3.260);
%    \node (5) [bound, fit=(3) (4), label={[align=center,rotate=90]left:\textbf{Salt loop}\\}] {};
%    \node (6) [bound, fit=(2) (5), label={[align=center,rotate=90]left:\textbf{Inner iteration}\\}] {};
%    \node (7) [bound, fit=(1) (6), label={[align=center,rotate=90]left:\textbf{Outer iteration}\\}] {};
%  \end{tikzpicture}
%  \caption{The nested iteration structure coupling the $S_N$, neutron diffusion,
%  in-core \gls{DNP}-temperature, and ex-core
%  solvers for the reactivity insertion simulation using the hybrid $S_N$-diffusion method.}
%  \label{fig:insertion-coupling}
%\end{figure}
%
%The reactivity insertion simulation requires a prior ``steady-state'' simulation to obtain the group
%flux, \gls{DNP}, and temperature distributions under 1 MW power operation and steady salt flow. The
%``steady-state'' simulation is a time-dependent simulation with timestep sizes of 1 s to advance
%the model towards 1 MW steady state operation. Following this, the reactivity insertion simulation
%uses the \gls{MOOSE} Restart capability to load in variable distributions for the initial conditions.
%
%\subsection{Preliminary Rod Worth Calculations}
%
%The control rod worth of the \gls{MSRE} quarter-core model with $^{233}$U fuel must be determined
%to estimate the initial rod height and rod travel necessary for the reactivity insertion
%simulation. We start this process with comparisons of Rod 1 worths between a full-core OpenMC model
%and experimental data. Table \ref{table:u233-rod-worth} compares the total rod worth of Rod 1 at
%922 K between the OpenMC model from this work
%and \gls{MSRE} data. The rod worth estimate from OpenMC agrees closely with the predicted (i.e.,
%calculated) \gls{MSRE} value. The measured \gls{MSRE} rod worth value is about 7 \% smaller than
%the predicted value. \gls{MSRE} researchers noted poor precision in their rod worth measurements
%due to ``wear and aging'' in the control rod drive assembly \cite{engel_zero-power_1972}. This may
%be responsible for some uncertainty. However, the actual rod worth is likely closer to the measured
%value than the predicted value, and the \gls{MSRE} model of this work is likely to contain as yet
%unknown biases. Nevertheless, the rod worth estimate is close enough that it should not affect the
%\gls{MSRE} power response following a fixed reactivity insertion.
%%
%\begin{table}[t]
%  \centering
%  \caption{Total rod worth of Rod 1 at 922 K from OpenMC full-core \gls{MSRE} model and \gls{MSRE}
